---
title: "Application of AKSA methodology on example"
output: 
  html_document:
    code_folding: show
    theme:
      base_font:
        google: Prompt
      heading_font:
        google: Proza Libre
---

```{r setup, include=FALSE}
if (requireNamespace("thematic")) 
  thematic::thematic_rmd(font = "auto")
knitr::opts_chunk$set(warning = FALSE, message = FALSE, fig.showtext = TRUE, comment = NA) 

```

```{r github AKSA, include=FALSE}
##Clone github repo from AKSA - Stat in med paper
#install.packages("usethis")
#usethis::create_from_github("aspapercode/predVal", protocol = "https", rstudio = FALSE,
#                            destdir = "/home/JAGE_CB/BOOK/predictive_chapter")


##Important function for AKSA evaluation
```

```{r}
##Package to install 
library(ggplot2)
library(ggpubr)
library(readxl)
library(dplyr)
library(brglm2)
library(detectseparation)
```

## Data import

In order to illustrate methodology application, We'll use data from POPLAR study.[Gandara, et al](https://www.nature.com/articles/s41591-018-0134-3) explored the predictiveness of blood TMB (bTMB) in atezolisumab vs docetaxel.  

```{r}
data_poplar <- read_excel("../data/41591_2018_134_MOESM3_ESM.xlsx", 
                       sheet = "POPLAR_Clinical_Data")

## keep only available data // this exploration if for illustration purpose only
## I do not recommend to evaluate BMK predictivness on complete data only /*\

data_poplar <- data_poplar %>% 
#filter NA data 
  filter(BCOR != ".") %>% 
  filter(TRT01P != ".") %>% 
  filter(btmb != ".") %>% 
  filter(BCOR != "NE") %>% 
  mutate(btmb = as.numeric(btmb)) %>% 
#combute combined answer by merging PD and SD patients
  mutate(OR = case_when(
  BCOR == "PD" ~ "PD+SD",
  BCOR == "SD" ~ "PD+SD",
  BCOR == "PR" ~ "PR")) %>%
#Create a binary end-point for logistic regression // 1 = responder
  mutate(ORbin = case_when(
  BCOR == "PD" ~ 0,
  BCOR == "SD" ~ 0,
  BCOR == "PR" ~ 1)
)

summary(data_poplar)
```


## eTMB according to traitement and response

Objective is to explore bTMB predictiveness: evaluate if the probability of being a responder is significantly associated to bTMB according to treatment arm. 


```{r}

##bTMB according to BCOR
ggboxplot(data_poplar, "BCOR", "btmb", fill = "TRT01P",
          color = "TRT01P", palette =c("#00AFBB", "#E7B800"),
          add = "jitter")+
            ggtitle("bTMB according to treatment and response BCOR")

##bTMB according to combined BCOR
ggboxplot(data_poplar, "OR", "btmb", fill = "TRT01P",
          color = "TRT01P", palette =c("#00AFBB", "#E7B800"),
          add = "jitter")+
            ggtitle("bTMB according to treatment and response combined BCOR")



```


## Probability of being a PR per treatment arm

```{r}

##fit a logistic regression based on intercation 
modelfull <- glm(ORbin ~ btmb + TRT01P + btmb*TRT01P, family = "binomial", data = data_poplar)
summary(modelfull)$coef

##draw the 2 curves + IC
logfit <- ggplot(data = data_poplar, aes(btmb, ORbin, fill = TRT01P, color = TRT01P)) +
  geom_point(alpha = .15) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"),  fullrange = TRUE) +
  ggtitle("Logistic regression model fit") +
  xlab("bTMB") +
  ylab("Probability of being PR") +
  scale_fill_manual(values = c("darkcyan", "goldenrod")) +
    scale_color_manual(values = c("darkcyan", "goldenrod")) +
    theme_bw() 

logfit

ggsave(filename = "logfit_poplar.png", plot = logfit, width = 8, height = 6)

```



```{r}

###AKSA illustration:

    # Obtain predicted values and se for 100 plausible bTMB for each treatment arm 
    btmb <- seq(min(data_poplar$btmb), max(data_poplar$btmb), length.out = 100)
    pred1 <- predict(modelfull, newdata = data.frame(btmb, TRT01P = "Docetaxel"), se.fit = TRUE)
    pred2 <- predict(modelfull, newdata = data.frame(btmb, TRT01P = "MPDL3280A"), se.fit = TRUE)

    # Calculate the contrast and its confidence interval
    contrast <- pred2$fit - pred1$fit
    se.constrat <- sqrt(pred1$se.fit^2 + pred2$se.fit^2)
    lwr.contrast <- contrast - qnorm(1 - 0.025) * se.constrat
    upr.contrast <- contrast + qnorm(1 - 0.025) * se.constrat
    
    # Randomly select couple of points in real value of bTMB
    max_contrast <- sample(1:length(btmb), 10000, replace=T)
    min_contrast <- sample(1:length(btmb), 10000, replace=T)
    
    # Compute contrast difference between the lowest and highest value of each couple 
    diff <- contrast[pmax(min_contrast, max_contrast)] - contrast[pmin(min_contrast, max_contrast)]
    se.diff <- sqrt(se.constrat[min_contrast]^2 + se.constrat[max_contrast]^2)
    
    # Select values in each difference distribution
    y <- rnorm(10000, mean=diff, sd=se.diff)

    # Prob that the Diff bw the contrasts is above 0
    mean(y > 0)
    
    # Density of differences

    dens <- density(y)
    dens_df <- data.frame(x = dens$x, y = dens$y)
    
    ggplot() +
      geom_area(data = filter(dens_df, x < 0), aes(x = x, y = y), fill = "darkcyan", alpha = 0.5) +  # à gauche de 0
      geom_area(data = filter(dens_df, x >= 0), aes(x = x, y = y), fill = "goldenrod", alpha = 0.5) +  # à droite de 0
      #geom_line(data = dens_df, aes(x = x, y = y), color = "darkcyan") +  # courbe de densité
      geom_vline(xintercept = 0, color = "goldenrod", lwd = 1, lty = 2) +  # ligne verticale
      labs(x = "y", y = "density") +
      theme_bw()
    

```


Then, we can evaluate the probability of $\Delta$ to be over any marked difference :

```{r}
   #Plot de proba 
  
    proba_md <- 1:100
    cutoff <- seq(from = -4, to = 4, length.out = 100)
      
    for (i in 1:100){proba_md[i] <-  mean(y > cutoff[i]) }
    
    proba <- data.frame(prob = proba_md, cutoff = cutoff)
    
    ggplot(proba) +
      geom_line(aes(x = cutoff, y = prob), color = "darkcyan", lwd = 2) +
      labs(y= "Probability", 
           x = "test")+
      theme(axis.title=element_text(size=10)) +
      geom_vline(xintercept = 0, color = "goldenrod", lwd = 1, lty=2) +
      geom_vline(xintercept = 0.4, color = "goldenrod", lwd = 1, lty=2) +
      geom_vline(xintercept = 1, color = "goldenrod", lwd = 1, lty=2)
      

```



```{r}
#library(ggplot2)
library(cowplot)

##theme for plots 

theme_publication <- theme_bw() +
  theme(
    # Titres
    plot.title = element_text(size = 9, hjust = 0.5),
    axis.title = element_text(size = 8, ),
    axis.text  = element_text(size = 8),
    
    # Légende
    legend.title = element_text(size = 7),
    legend.text  = element_text(size = 6),
    legend.position = "right",   # ou "bottom" si tu veux compacte
    legend.key.size = unit(0.6, "cm"), # réduit la taille des symboles
    
    # Marges
    plot.margin = margin(3, 3, 3, 3)
  )

   #Plot for predictive chapter 

  ##Logistic fit and observed value on poplar study 

logfit_bw <-ggplot(data = data_poplar, aes(btmb, ORbin, fill = TRT01P, color = TRT01P)) +
  geom_point(alpha = .15) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"),  fullrange = TRUE) +
  ggtitle("Probability of response according to bTMB") +
  xlab("bTMB") +
  ylab("Probability of being PR") +
  scale_fill_manual(values = c("grey", "black")) +
  scale_color_manual(values = c("grey", "black")) +
  theme_publication


  ##Density of differences and probability of difference being above or below zero

p_left  <- 0.15   # 
p_right <- 0.85   # 

dens_df <- dens_df %>%
  mutate(region = ifelse(x < 0, "left", "right"))
  
  
diffdensity <- ggplot() +
  geom_area(data = filter(dens_df, x < 0), 
            aes(x = x, y = y, fill = "left"), alpha = 0.5) +
  geom_area(data = filter(dens_df, x >= 0), 
            aes(x = x, y = y, fill = "right"), alpha = 0.5) +
  geom_vline(xintercept = 0, color = "black", lwd = 0.8, lty = 2) +
    ggtitle("Distribution of the difference") +
  labs(x = "y", y = "density") +
  scale_fill_manual(
    values = c("left" = "grey70", "right" = "black"),
    labels = c(
      bquote(P(Delta < 0) == .(round(p_left, 3))),
      bquote(P(Delta >= 0) == .(round(p_right, 3)))
    ),
    name = "Probability"
  ) +
  xlab(expression(Delta)) +
 theme_publication


##probability plot:
vlines <- data.frame(
  cutoff = c(0, 1, 2),
  label  = c("0", "1", "2")
)

  probaplot <- 
ggplot(proba) +
  geom_line(aes(x = cutoff, y = prob), 
            color = "black", lwd = 1) +
  labs(y = "Probability", x = "thresholds") +
#  theme(axis.title = element_text(size = 10)) +
  geom_vline(data = vlines, 
             aes(xintercept = cutoff, linetype = label), 
             color = "black", lwd = 0.8) +
  scale_linetype_manual(
    values = c("dashed", "dotdash", "dotted"),
    labels = c(
      expression(P(Delta  >=  0) == 0.85),
      expression(P(Delta  >=  1) == 0.7),
      expression(P(Delta  >=  2) == 0.53)
    )
  ) +
     ggtitle(expression("Probability that " * Delta * "  exceeds given thresholds")) +
  theme_publication +
  labs(linetype = "Probability thresholds")

# Combining 3 plots
#combined_plot <- plot_grid(logfit_bw, diffdensity, probaplot,
#                           labels = c("A", "B", "C"),   
#                           ncol = 2, align = "h")       

combined_plot1 <- plot_grid(diffdensity, probaplot, 
                           labels = c("B", "C"),   # ajoute A, B, C
                           ncol = 2, align = "h")       # côte à côte

combined_plot2 <- plot_grid(logfit_bw, combined_plot1,
                           labels = c("A", ""),   # ajoute A, B, C
                           nrow = 2, align = "h")       # côte à côte
# Titre global
title <- ggdraw() + 
  draw_label("AKSA application on POPLAR database", fontface = 'bold', x = 0.5, hjust = 0.5, size = 10)

final_plot <- plot_grid(title, combined_plot2, ncol = 1, rel_heights = c(0.1, 1))


# Affichage
final_plot

# Sauvegarde
ggsave("combined_plot.png", final_plot, width = 8, height = 8, dpi = 300)


```

